\text{Additive drift with tail bounds}

### 2021-11-28 article.tex
some changes in graphics and adding new plot

\newcommand\setItemnumber[1]{\setcounter{enumi}{\numexpr#1-1\relax}}

@@

\newcommand{\expx}[1]{e^{-|x|^{#1}}}
\newcommand{\expxpoz}[1]{e^{-x^{#1}}}
\newcommand{\infint}[1]{\int_{-\infty}^{+\infty} #1 \, dx}
\newcommand{\infintpoz}[1]{\int_{0}^{+\infty} #1 \, dx}
\newcommand{\der}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\cm}{coming soon\dots}

@@

>>>>

\newcommand\setItemnumber[1]{\setcounter{enumi}{\numexpr#1-1\relax}}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

\newcommand{\expx}[1]{e^{-|x|^{#1}}}
\newcommand{\expxpoz}[1]{e^{-x^{#1}}}
\newcommand{\infint}[1]{\int_{-\infty}^{+\infty} #1 \, dx}
\newcommand{\infintpoz}[1]{\int_{0}^{+\infty} #1 \, dx}
\newcommand{\der}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\cm}{coming soon\dots}
\newcommand{\expp}[1]{E[e^{#1 \Delta_t}]}

@@

\subsection{Comparing results for processes with negative drifts}

Consider a process \( \{ X_n \}_{n \in \mathbb{N}}\) with negative drift (i.e. \(E[X_{t + 1} - X_{t} | X_t = s] < 0\) for all \(s\)) and special requirement for probability:

\[
    \Pr[X_{t + 1} - X_{t} = j] \leq \frac{r}{(1 + \varepsilon)^{j + s}},
\]

for some \(\varepsilon, r, s > 0\) and for all \(j \geq -s\).

Because this process has negative drift we need to impose restrictions for values of \(s\). Also because summation of probabilities for all possible values of \(X_{t + 1} - X_{t}\) must be equals 1 we need to find \(r\).

\begin{align*}
    \sum_{j = -s}^{+\infty} Pr[X_{t + 1} - X_t = j] &\leq \sum_{j = -s}^{+\infty} \frac{r}{(1 + \varepsilon)^{j + s}} = r \sum_{j = 0}^{+\infty} \frac{1}{(1 + \varepsilon)^{j}} \\
    1 &\leq r \frac{1}{1 - \frac{1}{1 + \varepsilon}} = r \frac{1 + \varepsilon}{\varepsilon} \\
    r &\geq \frac{\varepsilon}{1 + \varepsilon} 
\end{align*}

Consider \(\Delta_t = X_{t + 1} - X_{t}\), than compute \(E[\Delta_t]\):

\begin{align*}
    E[\Delta_t] &= \sum_{j = -s}^{+\infty} j Pr[\Delta_t = j] \leq \sum_{j = -s}^{+\infty} j \frac{r}{(1 + \varepsilon)^{j + s}} \leq \sum_{j = 0}^{+\infty} (j - s) \frac{\varepsilon}{(1 + \varepsilon)^{j + 1}} = \\
    &= \sum_{j = 0}^{+\infty} j \frac{\varepsilon}{(1 + \varepsilon)^{j + 1}} - s \sum_{j = 0}^{+\infty}\frac{\varepsilon}{(1 + \varepsilon)^{j + 1}} = \frac{1}{\varepsilon} - s
\end{align*}

Also compute \(E\left[e^{\gamma \Delta_t}\right]\) for all positive \(\gamma\):

\begin{align*}
    E[e^{\gamma \Delta_t}] &= \sum_{j = -s}^{+\infty} e^{\gamma j} \Pr[\Delta_t = j] \leq e^{-\gamma s} \sum_{j = 0}^{+\infty} e^{\gamma j} \frac{\varepsilon}{(1 + \varepsilon)^{j + 1}} = \\
    &= e^{-\gamma s} \frac{\varepsilon}{1 + \varepsilon} \sum_{j = 0}^{+\infty} \left(\frac{e^\gamma}{1 + \varepsilon}\right)^j = e^{-\gamma s} \frac{\varepsilon}{1 + \varepsilon} \frac{1}{(1 - \frac{e^\gamma}{1 + \varepsilon})} = \frac{\varepsilon e^{-\gamma s}}{(1 + \varepsilon) - e^\gamma} 
\end{align*}

Finding minimum of previous function can be useful, so lets find \(\gamma_0\) where \(\frac{\partial E[e^{\gamma \Delta_t}]}{\partial \gamma} = 0\):

\begin{align*}
    \frac{\partial E[e^{\gamma \Delta_t}]}{\partial \gamma} &= \frac{\partial }{\partial \gamma} \left(\frac{\varepsilon e^{-\gamma s}}{(1 + \varepsilon) - e^\gamma}\right) = \varepsilon \frac{-s e^{-\gamma s} (1 + \varepsilon - e^\gamma) - e^{-\gamma s}(-e^\gamma)}{(1 + \varepsilon - e^\gamma)^2} = \\
    &= \frac{-\varepsilon e^{-\gamma s}}{(1 + \varepsilon - e^\gamma)^2} (e^\gamma (s + 1) - s (1 + \varepsilon)) = 0
\end{align*}
\begin{align*}
    e^{\gamma_0} (s + 1) - s (1 + \varepsilon) = 0
\end{align*}
\begin{align*}
    \gamma_0 = \ln\left(\frac{s}{s + 1}(1 + \varepsilon)\right)
\end{align*}

Define some variables for next calculations:

\begin{itemize}
    \item \(\gamma_1\) - point where math expectation of exponent intersects 1:
    \begin{align*}
        E[e^{\gamma_1 \Delta_t}] = 1
    \end{align*}
    \item \(\delta_0\) - point where math expectation of exponent intersects sub-gaussian boundary:
    \begin{align*}
        E[e^{\delta_0 \Delta_t}] = e^{\frac{c}{2}\delta_0^2}
    \end{align*}
    Of course if \(c = 0\) then \(\delta_0 = \gamma_1\)
    \item \(\gamma_m\) - point where probability by negative drift reach minimum
    \begin{align*}
        \gamma_m = \argmin \left(\ \frac{d}{s} \cdot p(\gamma) \cdot e^{-\gamma d}\ \right),
    \end{align*}
    where \(d := \frac{b - a}{2}\)

    Because we want probability to be as small as possible, define \(p(\gamma) = \frac{1}{1 - E[e^{\gamma \Delta_t}]}\). Consider derivative of probability function to conclude about boundaries of \(\gamma_m\):

    \begin{align*}
        \der{}{\gamma} \left(\frac{d e^{-\gamma d}}{s} p \right)= \frac{d e^{-\gamma d}}{s} \left(p'_\gamma - dp\right) = \frac{d e^{-\gamma d}}{s (1 - E[e^{\gamma \Delta_t}])} \left(\frac{E'_\gamma [e^{\gamma \Delta_t}]}{1 - E[e^{\gamma \Delta_t}]} - d\right)
    \end{align*}

    For all \(\gamma \in (0, \gamma_0)\) we have: \(\expp{\gamma} \in (0, 1)\) and \(E'_\gamma[e^{\gamma \Delta_t}] < 0\), hence derivative of probability is negative for all \(d\). After it for all \(\gamma \in [\gamma_0, \gamma_1)\) we have same limits for \(\expp{\gamma}\) and, because \(E'_\gamma[e^{\gamma \Delta_t}] \geq 0\), it grows to 1, hence somewhere between \(\gamma_0\) and \(\gamma_1\) derivative becomes 0, hence

    \begin{align*}
        \gamma_m \in [\gamma_0, \gamma_1)
    \end{align*}

\end{itemize}
